Ethnicity sensitive author disambiguation of semi-labeled corpora of scientific articles

### 2015-08-24 paper/paper.tex
structure

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem,
digital libraries are increasingly offering tools for authors to manually
curate their publications and claim which ones are theirs. Indirectly, these
tools allow for the inexpensive collection of large annotated training data,
which can be further leveraged to build a complementary automated
disambiguation system capable of inferring patterns for identifying
publications written by a same person.  Building upon more than 1 million
crowdsourced annotations, we propose an automated author disambiguation
solution exploiting this data (i) to learn an accurate classifier for
identifying corefering authors and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way. To the best of our
knowledge, our analysis is the first to be carried out on data of this size and
coverage. It reveals that... \glnote{To be completed with actual results.}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...).
Recently, efforts also started to create persistent digital identifiers
assigned to researchers (e.g., ORCID, ResearchedID, ...), with the further
objective to embed these identifiers in the submission workflow of publishers
or repositories (e.g., Elsevier, arXiv, Inspire-HEP, ...), thereby univocally
solving any disambiguation issue. With the large cost of centralized manual
authorship control, or until the larger adoption of crowdsourced solutions, the
impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of
persistent digital identifier efforts is conditioned to a large and ubiquitous
adoption by both researchers and publishers. For these reasons, fully automated
machine learning-based methods have been proposed during the past decade to
provide immediate, less costly and efficient solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.  In particular, we study
how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying corefering authors and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way. Our analysis of parameters
and features on this large training set reveals that...

\glnote{To be completed with actual results.}\\
\glnote{Also mention error analysis, suggesting that
uncertain disambiguation results should be suggested to human curators, thereby
closing the loop and creating a virtuous circle.}\\
\glnote{New idea: phonetization in blocking}\\
\glnote{New idea: ethnicity features}

>>>>

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem,
digital libraries are increasingly offering tools for authors to manually
curate their publications and claim which ones are theirs. Indirectly, these
tools allow for the inexpensive collection of large annotated training data,
which can be further leveraged to build a complementary automated
disambiguation system capable of inferring patterns for identifying
publications written by a same person.  Building upon more than 1 million of
publicly released crowdsourced annotations, we propose an automated author
disambiguation solution exploiting this data (i) to learn an accurate
classifier for identifying corefering authors and (ii) to guide the clustering
of scientific publications by distinct authors in a semi-supervised way. To the
best of our knowledge, our analysis is the first to be carried out on data of
this size and coverage. With respect to the state of the art, we validate the
general pipeline used in most existing solutions, and improve by proposing
phonetic-based blocking strategies, thereby increasing recall, and strong
ethnicity-sensitive features for learning a linkage function, thereby tailoring
disambiguation to non-Western author names whenever necessary.

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...).
Recently, efforts also started to create persistent digital identifiers
assigned to researchers (e.g., ORCID, ResearchedID, ...), with the further
objective to embed these identifiers in the submission workflow of publishers
or repositories (e.g., Elsevier, arXiv, Inspire-HEP, ...), thereby univocally
solving any disambiguation issue. With the large cost of centralized manual
authorship control, or until the larger adoption of crowdsourced solutions, the
impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of
persistent digital identifier efforts is conditioned to a large and ubiquitous
adoption by both researchers and publishers. For these reasons, fully automated
machine learning-based methods have been proposed during the past decade to
provide immediate, less costly and efficient solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.  In particular, we study
how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying corefering authors and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way. Our analysis of parameters
and features on this large dataset reveals that the general pipeline
commonly used in existing solutions is an effective solution for author disambiguation.
In addition, and  as validated
by experimental results, we propose alternative strategies for blocking based on the
phonetization of author names, thereby increasing recall. We also propose
effective ethnicity-sensitive features for learning a linkage function,
thereby tailoring author
disambiguation to non-Western author names whenever necessary.
